Follow the Flow of Rain and Paint: \break An Output Sensitive Contour Tree Algorithm

### 2014-09-22 paint-contour-tree.tex
more changes

\begin{definition} \label{def:support} A light path $p$ is called \emph{supported} if there exists
a contiguous (from $p$) chain of ancestors $q_1, q_2, \ldots, q_k$ (in $P(T)$) such that $|H_{q_k}| \leq 99|H_p|/100$
and all $q_i$'s are heavy.
\end{definition}

@@

We now construct the shrub of unsupported light paths. Consider $p \in \cL \setminus \cL'$,
and traverse the chain of ancestors from $p$. Eventually, we must reach another light path $q$.
(If not, we have reached the root $r$ of $\pmax(T)$. But $|H_r| = 0$, implying $p$ is supported.)
Insert edge from $p$ to $q$, so $q$ is the parent of $p$ in $\cU$. This construction leads
the shrub forest of $\cL \setminus \cL'$, where every unsupported light path is an internal node, and all
roots are supported light paths. \Lem{support} is a direct consequence of the following,
which is the really the main lemma.

\begin{lemma} \label{lem:cu} Let $\cU$ denote a connected component (shrub) in the shrub forest of $\cL \setminus \cL'$ and let $r$
be the root of $\cU$. Then $\sum_{p \in \cU} \sum_{v \in p} \log |H_v| = O(|H_r| \log |H_r|)$.
\end{lemma}

We proceed with the setup to prove \Lem{cu}.
We focus our attention on some shrub $\cU$.
We prove a technical lemma about $\cU$, from which \Lem{cu} follows directly.

\begin{lemma} \label{lem:cu-root} (i) For any $v \in p$ such that $p \in \cU$,
$|H_v| = O(|H_r|)$. (ii) $\sum_{p \in \cU} |p| = O(|H_r|)$.
\end{lemma}

\begin{proof} (of \Lem{cu}) We have $\sum_{p \in \cU} \sum_{v \in p} \log |H_v| = O((\log |H_r|) \sum_{p \in \cU} |p|) = O(|H_r|\log |H_r|)$.
\end{proof}

\subsection{The root is everything in $\cU$} \label{sec:weight}

@@

>>>>

\begin{definition} \label{def:support} A light path $p$ is called \emph{supported} if: 
at least $|H_p|/100$ vertices in $p$ lie in heavy ancestors $q$ such that $|H_q| > |H_p|/100$.
%
%
% there exists
%a contiguous (from $p$) chain of ancestors $q_1, q_2, \ldots, q_k$ (in $P(T)$) such that $|H_{q_k}| \leq 99|H_p|/100$
%and all $q_i$'s are heavy.
\end{definition}

@@

We now construct the shrub of unsupported light paths. Consider $p \in \cL \setminus \cL'$,
and traverse the chain of ancestors from $p$. Eventually, we must reach another light path $q$.
(If not, we have reached the root $r$ of $\pmax(T)$. But $|H_r| = 0$, implying $p$ is supported.)
Insert edge from $p$ to $q$, so $q$ is the parent of $p$ in $\cU$. This construction leads
to the shrub forest of $\cL \setminus \cL'$, where every unsupported light path is an internal node, and all
roots are supported light paths. \Lem{support} is a direct consequence of the following,
which is the really the main lemma.

%\begin{lemma} \label{lem:cu} Let $\cU$ denote a connected component (shrub) in the shrub forest of $\cL \setminus \cL'$ and let $r$
%be the root of $\cU$. Then $\sum_{p \in \cU} \sum_{v \in p} \log |H_v| = O(|H_r| \log |H_r|)$.
%\end{lemma}
%
%We proceed with the setup to prove \Lem{cu}.
%We focus our attention on some shrub $\cU$.
%We prove a technical lemma about $\cU$, from which \Lem{cu} follows directly.

\begin{lemma} \label{lem:cu-root} Let $\cU$ denote a connected component (shrub) in the shrub forest of $\cL \setminus \cL'$ and let $r$
be the root of $\cU$. (i) For any $v \in p$ such that $p \in \cU$,
$|H_v| = O(|H_r|)$. (ii) $\sum_{p \in \cU} |p| = O(|H_r|)$.
\end{lemma}

\begin{proof} (of \Lem{support}) We have $\sum_{p \in \cU} \sum_{v \in p} \log |H_v| = O((\log |H_r|) \sum_{p \in \cU} |p|) = O(|H_r|\log |H_r|)$.
We sum this over all $\cU$ in the shrub forest, and note that roots in the shrub forest are supported light paths.
\end{proof}

\subsection{Proving \Lem{cu-root}: the root is everything in $\cU$} \label{sec:weight}

@@

This immediately proves part (i) of \Lem{cu-root}. Part (ii) requires some more work.

\begin{proof} (part (ii) of \Lem{cu-root}) Let level $\ell$ of $\cU$ denote all nodes at distance $\ell$ from the root $r$.
These set of these nodes is designated as $\cU_\ell$. We set $\sigma_\ell = \sum_{p \in \cU_\ell} |p|$
and $\sigma_{\leq \ell} = \sum_{k \leq \ell} \sigma_k$.
By \Lem{geometric}, for any $p \in \cU_\ell$, $|H_p| \leq (2/3)^\ell |H_r|$.

Since all nodes in $\cU$ represent light paths, $\forall p \in \cU$, $|p| < \sqrt{|H_p|}/100$.
So $\sigma_\ell < \sum_{p \in \cU_\ell} \sqrt{|H_p|}/100$.
